---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# vecplyr

<!-- badges: start -->
<!-- badges: end -->

`vecplyr` extends 2 dplyr functions to vectors: `filter` and `mutate`. They are inspired in the `purrr` functions `modify` and `keep`/`discard`, but much more efficient since they are restricted to vectors and vectorized functions.

## Installation

You can install the development version of vecplyr from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("vituri/vecplyr")
```

## Usage

`vecplyr` has two main functions: 

- `v_filter` throws away some entries of a vector based on a condition;
- `v_mutate` modifies a vector with a function when some condition is TRUE.

Suppose we want to do the following: given an integer vector `x`, we want to:

- keep only the even numbers of x;
- add 1 to all numbers of x;
- multiply by -1 the numbers less than 25.
- sum the result

```{r}
library(vecplyr)
x = 1:50

# vecplyr approach can be "piped"
x %>% 
  v_filter(\(x) x %% 2 == 0) %>% # keep the even
  v_mutate(\(x) x+1) %>% # sum 1
  v_mutate(\(x) -x, \(x) x < 25) %>% # multiply by -1 only those that satisfy x < 25 
  sum()

# or using purrr formulas
x %>% 
  v_filter(~ .x %% 2 == 0) %>% # keep the even
  v_mutate(~ .x+1) %>% # sum 1
  v_mutate(~ -.x, ~ .x < 25) %>% # multiply by -1 only those that satisfy x < 25
  sum()

# base R approach
y = x[x %% 2 == 0]
z = y + 1
w = z
id = w < 25
w[id] = - w[id]
sum(w)
```



## Examples

```{r example}
library(vecplyr)

x = 1:1e5

# get the sum of all odd numbers in x
x %>% v_filter(\(x) x %% 2 == 0) %>% sum()

# get all numbers where the square of it is less than 10000
x %>% v_filter(\(x) x^2 < 10000)

# take values from column b when a is NA
tibble::tibble(
  a = 1:10 %>% v_mutate(NA, \(x) x %% 2 == 0)
  ,b = -(1:10)
) %>% 
  dplyr::mutate(
    if_a_is_na_then_b = a %>% v_mutate(b, is.na)
  )

```

# Benchmark

Compared to `purrr` functions `keep` and `modify`, `vecplyr` is faster because it only accepts vectors and vectorized functions.

```{r}
x = 1:1e5

mbm = microbenchmark::microbenchmark(
  vecplyr = {
    x %>% v_filter(~ .x^2 <= 5000)
  }
  ,purrr = {
    x %>% purrr::keep(~ .x^2 <= 5000)
  }

  ,times = 25L
)

mbm
```

```{r plot1}
ggplot2::autoplot(mbm)
```

```{r}
x = 1:1e5

mbm2 = microbenchmark::microbenchmark(
  vecplyr = {
    x %>% v_mutate(~ .x + 2, ~ .x <= 10000)
  }
  ,purrr = {
    x %>% purrr::modify_if(.f = ~ .x + 2L, .p = ~.x <= 10000)
  }

  ,times = 25L
)

mbm2
```

```{r plot2}
ggplot2::autoplot(mbm2)
```
